---
output: github_document
title: Homework Week 01
author: Alec L. Robitaille
date: 2022-01-26
---

```{r}
#| include = FALSE

knitr::opts_chunk$set(
	fig.path = '../graphics/homework/'
)
```




# Question 1

> Suppose the globe tossing data had turned out to be 4 water and 11 land.
Construct the posterior distribution, using grid approximation. Use the same
flat prior as in the book.


```{r w01_q01}
# Packages --------------------------------------------------------------------
library(data.table)


# Process ---------------------------------------------------------------------
# Size of grid for grid approximation
gridsize <- 1000

# Prior grid
prior_grid <- seq(0, 1, length.out = gridsize)

# Prior probability (all 1)
prior_prob <- rep(1, gridsize)

# Data probability, using binomial distribution
water <- 4
land <- 11
data_prob <- dbinom(water, water + land, prob = prior_grid)

# Wrap as DT
DT_q01 <- data.table(prop_water = prior_grid,
                     prior_water = prior_prob,
                     prob_water = data_prob)

# Calculate the posterior numerator by multiplying prior and data probability
DT_q01[, posterior_num := prior_water * prob_water]

# Standardize by sum of posterior numerator
DT_q01[, posterior := posterior_num / sum(posterior_num)]


# Plot -------------------------------------------------------------------------
plot(DT_q01$prop_water, DT_q01$posterior, type = 'l')
```


## Question 2

>  Now suppose the data are 4 water and 2 land. Compute the posterior again, but
this time use a prior that is zero below p = 0.5 and a constant above p = 0.5.
This corresponds to prior information that a majority of the Earth’s surface is
water.


```{r w01_q02}
# Process ---------------------------------------------------------------------
# Size of grid for grid approximation
gridsize <- 1000

# Prior grid
prior_grid <- seq(0, 1, length.out = gridsize)

# Prior probability (all 1)
prior_prob <- c(rep(0, gridsize / 2), rep(1, gridsize / 2))

# Data probability
#  given 4, using binomial distribution
water <- 4
land <- 2
data_prob <- dbinom(water, water + land, prob = prior_grid)

# Wrap as DT
DT_q02 <- data.table(prop_water = prior_grid,
                     prior_water = prior_prob,
                     prob_water = data_prob)

# Calculate the posterior numerator by multiplying prior and data probability
DT_q02[, posterior_num := prior_water * prob_water]

# Standardize by sum of posterior numerator
DT_q02[, posterior := posterior_num / sum(posterior_num)]


# Plot -------------------------------------------------------------------------
plot(DT_q02$prop_water, DT_q02$posterior, type = 'l')
```


## Question 3

> For the posterior distribution from 2, compute 89% percentile and HPDI
intervals. Compare the widths of these intervals. Which is wider? Why? If you
had only the information in the interval, what might you misunderstand about the
shape of the posterior distribution?



```{r w01_q03}
# Packages ---------------------------------------------------------------------
library(rethinking)
library(ggplot2)


# Samples ----------------------------------------------------------------------
# Sample from posterior
samples <- DT_q02[, sample(prop_water, size = gridsize, prob = posterior, replace = TRUE)]
mean(samples)
PI(samples, 0.89)
HPDI(samples, 0.89)

DT_q03 <- data.table(rbind(t(PI(samples, 0.89)), t(HPDI(samples, 0.89))), c('PI', 'HDPI'))
setnames(DT_q03, c('low', 'high', 'type'))


# Plot -------------------------------------------------------------------------
ggplot() +
  geom_line(aes(prop_water, posterior), data = DT_q02) +
  geom_linerange(aes(xmin = low, xmax = high, y = 0.0033, color = 'PI'), 
                 size = 3, data = DT_q03[type == 'PI']) +
  geom_linerange(aes(xmin = low, xmax = high, y = 0.0035, color = 'HDPI'), 
                 size = 3, data = DT_q03[type == 'HDPI']) +
  scale_color_manual(values = list('PI' = 'green', 'HDPI' = 'purple'),
                     name = 'Interval type')
```

* PI: percentile intervals
* HDPI: highest posterior density interval 

Percentile intervals assign equal probability mass to each tail whereas 
highest posterior density intervals return the interval with the highest posterior
probability. 