---
output: 
  github_document:
    html_preview: false
title: Homework Week 02
author: Alec L. Robitaille
date: 2022-02-07
---


```{r}
#| include = FALSE
knitr::opts_knit$set(root.dir = '../.')
```

```{r}
#| include = FALSE
source('R/packages.R')
source('R/dag_plot.R')

knitr::opts_chunk$set(
	echo = FALSE,
	fig.path = '../graphics/homework/'
)


theme_set(theme_bw(base_size = 16))
```

# Question 1

> Construct a linear regression of weight as predicted by height, using the adults
(age 18 or greater) from the Howell1 dataset. The heights listed below were
recorded in the !Kung census, but weights were not recorded for these
individuals. Provide predicted weights and 89% compatibility intervals for each
of these individuals. That is, fill in the table below, using model-based
predictions.


```{r predict_h02_q01}
ggplot(tar_read(predict_h02_q01), 
       aes(x = factor(height), y = weight)) +
  labs(x = 'height', y = 'predicted weight') + 
  guides(fill = "none") + 
  stat_halfeye(aes(fill = stat(cut_cdf_qi(cdf, .width = c(.89, 1))))) + 
  scale_fill_grey(start = 0.4, end = 0.6)
```



# Question 2

> From the Howell1 dataset, consider only the people younger than 13 years old.
Estimate the causal association between age and weight. Assume that age
influences weight through two paths. First, age influences height, and height
influences weight. Second, age directly influences weight through age- related
changes in muscle growth and body proportions. All of this implies this causal
model (DAG):

```{r dag_h02_q02}
dag <- dagify(
    W ~ A + H,
    H ~ A,
    exposure = 'A',
    outcome = 'W'
)

dag_plot(dag)

print('Variables to include to measure total effect of age on weight:')
adjustmentSets(dag, effect = 'total')
```


> Use a linear regression to estimate the total (not just direct) causal effect
of each year of growth on weight. Be sure to carefully consider the priors. Try
using prior predictive simulation to assess what they imply.


```{r priors_h02_q02}
ggplot(tar_read(DT_priors_h02_q02), aes(age, weight)) + 
  xlim(0, 14) + ylim(0, 50) + 
  geom_abline(aes(intercept = age, slope = beta_age))
```

```{r h02_q02_summary}
tar_load(q02_summary_h02_q02)
setDT(q02_summary_h02_q02)
q02_summary_h02_q02[variable != 'lp__', .(variable, mean, q5, q95)]
```


# Question 3 

> Now suppose the causal association between age and weight might be different
for boys and girls. Use a single linear regression, with a categorical variable
for sex, to estimate the total causal effect of age on weight separately for
boys and girls. How do girls and boys differ? Provide one or more posterior
contrasts as a summary.



```{r h02_q03_contrast}
ggplot(tar_read(q03_contrast),
         aes(sim_age, predict_wt)) + 
    stat_halfeye() +
    scale_x_continuous(breaks = c(0, 5, 10), labels = c(0, 5, 10)) + 
    labs(x = 'age', y = 'posterior weight contrast', 
         title = 'Posterior predicted weight contrast (male - female)')
```