---
output: 
  github_document:
    # html_preview: false
title: Homework Week 03
author: Alec L. Robitaille
date: 2022-02-16
---


```{r}
#| include = FALSE
knitr::opts_knit$set(root.dir = '../.')
```

```{r}
#| include = FALSE
source('R/packages.R')
source('R/dag_plot.R')

knitr::opts_chunk$set(
	echo = FALSE,
	fig.path = '../graphics/homework/'
)


theme_set(theme_bw(base_size = 16))
```

# Question 1

> The first two problems are based on the same data. The data in data(foxes) are
116 foxes from 30 different urban groups in England. These fox groups are like
street gangs. Group size (groupsize) varies from 2 to 8 individuals. Each group
maintains its own (almost exclusive) urban territory. Some territories are
larger than others. The area variable encodes this information. Some territories
also have more avgfood than others. And food influences the weight of each fox.
Assume this DAG:

> where F is avgfood, G is groupsize, A is area, and W is weight. Use the
backdoor criterion and estimate the total causal influence of A on F. What
effect would increasing the area of a territory have on the amount of food
inside it?


# Question 2

> Now infer both the total and direct causal effects of adding food F to a
territory on the weight W of foxes. Which covariates do you need to adjust for
in each case? In light of your estimates from this problem and the previous one,
what do you think is going on with these foxes? Feel free to speculate— all that
matters is that you justify your speculation.


# Question 3 

> Reconsider the Table 2 Fallacy example (from Lecture 6), this time with an
unobserved confound U that influences both smoking S and stroke Y. Here’s the
modified DAG:

