---
output: 
  github_document:
    html_preview: false
title: Homework Week 05
author: Alec L. Robitaille
date: 2022-04-21
---

```{r}
#| include = FALSE
knitr::opts_knit$set(root.dir = '../.')
```

```{r}
#| include = FALSE
source('R/packages.R')
source('R/dag_plot.R')
source('R/plot_halfeye.R')

knitr::opts_chunk$set(
	echo = FALSE,
	fig.path = '../graphics/homework/'
)


theme_set(theme_bw(base_size = 16))
```


## Question 1

> The data in data(NWOGrants) are outcomes for scientific funding applications
for the Netherlands Organization for Scientific Research (NWO) from 2010–2012
(see van der Lee and Ellemers doi:10.1073/pnas.1510159112). These data have a
very similar structure to the UCBAdmit data discussed in Chapter 11. Draw a DAG
for this sample and then use one or more binomial GLMs to estimate the TOTAL
causal effect of gender on grant awards.

```{r dag_h05_q01}
dag <- dagify(
  Award ~ Gender + Discipline,
  Discipline ~ Gender,
  exposure = 'Gender',
  outcome = 'Award'
)
dag_plot(dag, 'graphics/homework/dot_h05_q01')
```

```{r model_h05_q01}
mod <- tar_read(h05_q01_brms_sample)

conditional_effects(mod)
```


## Question 2

> Now estimate the DIRECT causal effect of gender on grant awards. Compute the
average direct causal effect of gender, weighting each discipline in proportion
to the number of applications in the sample. Refer to the marginal effect
example in Lecture 9 for help.

```{r model_h05_q02}
mod2 <- tar_read(h05_q02_brms_sample)

conditional_effects(mod2, effects = 'gender')
conditional_effects(mod2, effects = 'discipline')
conditional_effects(mod2)

draws <- as_draws_df(mod2)
DT <- tar_read(DT_grants)

counts <- DT[, .(applications = sum(applications)), discipline]

DT_m <- copy(counts)[, gender := unique(DT$gender)[1]]
DT_f <- copy(counts)[, gender := unique(DT$gender)[2]]

new_data <- rbindlist(list(DT_m, DT_f))
pred <- data.table(add_predicted_draws(new_data, mod2))

ggplot(pred) + 
  stat_halfeye(aes(x = .prediction / applications, gender)) + 
  facet_wrap(~discipline)

x <- pred[gender == 'f', .prediction / applications] - 
  pred[gender == 'm', .prediction / applications]
ggplot(data.frame(x = x)) + 
  stat_halfeye(aes(x)) + 
  geom_vline(xintercept = 0, color = 'grey') + 
  labs(x = 'f - m', y = '')
```


## Question 3

> Considering the total effect (problem 1) and direct effect (problem 2) of
gender, what causes contribute to the average difference between women and men
in award rate in this sample? It is not necessary to say whether or not there is
evidence of discrimination. Simply explain how the direct effects you have
estimated make sense (or not) of the total effect.

Above the horizontal line indicate higher proportion of men were accepted,
below the line indicates the reverse. The color of the point indicates
the discipline specific, proportion of female applications.

```{r h05_q03}
DT[, total_apps := sum(applications), gender]
DT_f_rates <- DT[, .(award_rate = awards / applications, 
                     app_rate_gender = applications / total_apps), 
                 by = .(gender, discipline)][gender == 'f']

cast <- dcast(
  pred[, .(discipline, gender, award = .prediction / applications)],
  formula = discipline ~ gender,
  fun.aggregate = mean,
  value.var = 'award'
)

ggplot(cast[DT_f_rates, on = 'discipline']) + 
  geom_point(size = 6, color = 'black', shape = 21, 
             aes(f, m, fill = app_rate_gender)) + 
  geom_abline(slope = 1, intercept = 0) + 
  scale_fill_scico() +
  labs(fill = 'Proportion of\nfemale \napplications') + 
  coord_equal(xlim = c(0.1, 0.3), ylim = c(0.1, 0.3))
```

<!-- 
## Question 4
4-OPTIONAL CHALLENGE. The data in data(UFClefties) are the outcomes
of 205 Ultimate Fighting Championship (UFC) matches (see ?UFClefties for de-
tails). It is widely believed that left-handed fighters (aka “Southpaws”) have an ad-
vantage against right-handed fighters, and left-handed men are indeed over-rep-
resented among fighters (and fencers and tennis players) compared to the general
population. Estimate the average advantage, if any, that a left-handed fighter has
against right-handed fighters. Based upon your estimate, why do you think left-
handers are over-represented among UFC fighters?
-->