---
output: 
  github_document:
    # html_preview: false
title: Homework Week 06
author: Alec L. Robitaille
date: 2022-04-22
---

```{r}
#| include = FALSE
knitr::opts_knit$set(root.dir = '../.')
```

```{r}
#| include = FALSE
source('R/packages.R')
source('R/dag_plot.R')
source('R/plot_halfeye.R')

knitr::opts_chunk$set(
	echo = FALSE,
	fig.path = '../graphics/homework/'
)


theme_set(theme_bw(base_size = 16))
```


## Question 1
> Conduct a prior predictive simulation for the Reedfrog model. By this I mean
to simulate the prior distribution of tank survival probabilities αj. Start by
using this prior: 

> αj ∼ Normal( ̄α, σ)

>̄ α ∼ Normal(0, 1)

> σ ∼ Exponential(1) 

> Be sure to transform the αj values to the probability scale for plotting and
summary. How does increasing the width of the prior on σ change the prior
distribution of αj? You might try Exponential(10) and Exponential(0.1) for
example.


Note alpha j is shown on the probability scale. 

```{r h06_q01}
tar_read(h06_q01_sim)
```


## Question 2
> Revisit the Reedfrog survival data, data(reedfrogs). Start with the varying
effects model from the book and lecture. Then modify it to estimate the causal
effects of the treatment variables pred and size, including how size might
modify the effect of predation. An easy approach is to estimate an effect for
each combination of pred and size. Justify your model with a DAG of this
experiment.


```{r h06_q02}
```


## Question 3
> Now estimate the causal effect of density on survival. Consider whether pred
modifies the effect of density. There are several good ways to include density
in your Binomial GLM. You could treat it as a continuous regression variable
(possibly standardized). Or you could convert it to an ordered category (with
three levels). Compare the σ (tank standard deviation) posterior distribution
to σ from your model in Problem 2. How are they different? Why?

```{r h06_q03}
```
