---
output: 
  github_document:
    # html_preview: false
title: Homework Week 07
author: Alec L. Robitaille
date: 2022-04-25
---

```{r}
#| include = FALSE
knitr::opts_knit$set(root.dir = '../.')
```

```{r}
#| include = FALSE
source('R/packages.R')
source('R/dag_plot.R')
source('R/plot_halfeye.R')
source('R/plot_rvars.R')

knitr::opts_chunk$set(
	echo = FALSE,
	fig.path = '../graphics/homework/'
)


theme_set(theme_bw(base_size = 16))
```


## Question 1
> The data in data(bangladesh) are 1934 women from the 1989 Bangladesh Fertility
Survey. For each woman, we know which district she lived in, her number of
living.children, her age.centered, whether she lived in an urban center, and
finally whether or not she used contraception (use.contraception).

> In this first problem, I only want you to investigate the proportion of women
using contraception in each district. Use partial pooling (varying effects).
Then compare the varying effect estimates to the raw empirical proportion in
each district. Explain the differences between the estimates and the data.

> Note that district number 54 is absent in the data. This causes some problems
in indexing the parameters. The simplest fix is just to tell ulam manually how
long the vector should be, like this: vector[61]:a ~ normal(abar,sigma). Pay
special attention to district number 54’s estimate

```{r h07_q01}
tar_read(plot_district_contraception)
```

## Question 2
> First, draw a DAG that includes all five variables: (1) Uses contraception C
(2) Age A (3) Children K (4) Urban U (5) District D. You don’t have to be an
expert on fertility. But do think about which variables can influence which
other variables. Second, design an estimation strategy to identify both the
total and direct causal effects of living in an urban center on contraceptive
use. Those are your estimands.

> Consider causal relationships among the variables. Then use your DAG to
justify an adjustment set that will yield the estimate of the causal effect of
urban living on contraceptive use.

> Do not run a statistical model (yet). I just want you to try to design an
analysis. There is no firm right answer. Just apply the backdoor criterion and
rules of d-separation (the elemental confounds) correctly to the DAG you design.

```{r dag_h07_q02}
dag <- dagify(
  Use ~ Age + District + Urban + NumbKids,
  NumbKids ~ Age + Urban,
  NumbKids ~ Age,
  Urban ~ District,
  exposure = 'Urban',
  outcome = 'Use'
)
dag_plot(dag, 'graphics/homework/dot_h07_q02')

adjustmentSets(dag, effect = 'total')
adjustmentSets(dag, effect = 'direct')
```


## Question 3
> Now build one or more statistical models to estimate the total and the direct
causal effects of urban living on contraceptive use. Again include district as a
simple varying effect (as in problem 1) so that each district has its own
average contraceptive use. You may also want to stratify the effect of urban
living by district. If you do, think carefully about how to do this
statistically.


```{r plot_h07_q03}
tar_read(plot_compare_total_direct)
```

<!--
4-OPTIONAL CHALLENGE. Can you also go beyond the parameter estimates
from problem 3 and compute a marginal causal effect of urban living for each dis-
trict, using a standard age distribution? Use any population age distribution you
like. The important thing is to project the estimates from the sample of each dis-
trict (which does not have representative age distributions) to the population. If
you think the different districts should have different age distributions, that would
be even more interesting.
-->