---
title: ""
output: github_document
---

## Homework Week 3
```{r, eval = T, include = F}
library(rethinking)
library(dagitty)
data("foxes")
d <- foxes
```

**Question 1:** The first two problems are based on the same data. The data in `data (foxes)` are 116 foxes from 30 different urban groups in England. These fox groups are like street gangs. Group size (`groupsize`) varies from 2 to 8 individuals. Each group maintains its own (almost exclusive) urban territory. Some territories are larger than others. The `area` variable encodes this information. Some territories also have more `avgfood` than others. And food influences the `weight` of each fox, where F is `avgfood`, G is `groupsize`, A is `area`, and W is `weight` . Use the backdoor criterion and estimate the total causal influence of A on F. What effect would increasing the area of a territory have on the amount of food inside it?

```{r draw-dag, echo = F}
dag1 <- dagitty("dag {
                A -> F 
                F -> G -> W 
                F -> W }")
coordinates(dag1) <- list(x = c(F = 1, A = 2, W = 2, G = 3), 
                               y = c(F = 0, A = -1, W = 1, G = 0))
drawdag(dag1)
```

Start off by identifying backdoor pathways and the potential adjustments that need to be made to the model to close backdoor paths:
```{r answer-1 dag}
dag1 <- dagitty("dag {
                A -> F 
                F -> G -> W 
                F -> W }")
adjustmentSets(dag1, exposure = "A", outcome = "F")
```

There are no backdoor pathways of A -> F (as seen in the DAG). So building a model with F as a function of A is sufficient: 
```{r answer-1 model}
# standardize variables 
d$F <- standardize(d$avgfood)
d$A <- standardize(d$area)
# model
m1 <- quap(
  alist(
    F ~ dnorm(mu, sigma),
    mu <- a + bA*A,
    a ~ dnorm(0,0.2),
    bA ~ dlnorm(0,0.5), # area can only be positive so constrain to log normal
    sigma ~ dexp(1) 
  ), data = d
)

precis(m1)
```

```{r answer-1 figure, echo = F}
# posterior predictions 
A_seq <- seq( from= -3 , to=3 , length.out=30 )
mu <- link( m1 , data=list(A=A_seq) )
mu.mean <- apply( mu , 2, mean )
mu.PI <- apply( mu , 2 , PI )
# plot it all
plot( F ~ A , data=d , col=rangi2 )
lines( A_seq , mu.mean , lwd=2 )
shade(mu.PI, A_seq )
```

The model output suggests that increasing the area territory would increase the amount of food available. 