---
title: ""
output: github_document
---

## Homework Week 4
```{r, eval = T, include = F}
library(rethinking)
library(brms)
library(dagitty)
```

**Question 1:** Revisit the marriage, age, and happiness collider bias example from Chapter 6. Run models m6.9 and m6.10 again (pages 178–179). Compare these two models using both PSIS and WAIC. Which model is expected to make better predictions, according to these criteria? On the basis of the causal model, how should you interpret the parameter estimates from the model preferred by PSIS and WAIC?

```{r}
d <- sim_happiness()
d2 <- d[d$age > 17, ]
d2$A <- (d2$age - 18) / (65 - 18) # rescale so age is one unit 
d2$mid <- d2$married + 1 # recode categorical variable 

m6.9 <- quap(
  alist(
    happiness ~ dnorm(mu, sigma),
    mu <- a[mid] + bA*A,
    a[mid] ~ dnorm(0,1),
    bA ~ dnorm(0,2),
    sigma ~ dexp(1)
  ), data = d2
)

m6.10 <- quap(
  alist(
    happiness ~ dnorm(mu, sigma),
    mu <- a + bA*A,
    a ~ dnorm(0,1),
    bA ~ dnorm(0,2),
    sigma ~ dexp(1)
  ), data = d2
)

compare(m6.9, m6.10, func=WAIC)

compare(m6.9, m6.10, func=PSIS)
```

Both WAIC and PSIS criterion rank m6.9 higher than m6.10, indicating that the model with both age and marriage status is the better predictor. However, in m6.9 we see that age has a negative relationship with happiness, because of the collider bias that is introduced by including both marriage status and age in the model. Although causally, m6.9 does not make sense or provide good inference, including both variables improves prediction enough to rank higher even when being "punished" for more variables.  

**Question 2:** Reconsider the urban fox analysis from last week’s homework. On the basis of PSIS and WAIC scores, which combination of variables best predicts body weight (W, weight)? How would you interpret the estimates from the best scoring model?  

**Coworking Group Question:** Take a question/model from your own work, and draw the DAG. Try to think about unobserved variables in addition to things you did measure. List the variables in your DAG and define them. What are the units? What is a reasonable range of values (using information from a different source than your own data)? Define or justify the edges in your DAG. Just a sentence or two, but enough to clarify why you/literature/science/logic think this relationship exists.  

**Question 3:** Build a predictive model of the relationship show on the cover of the book, the relationship between the timing of cherry blossoms and March temperature in the same year. The data are found in data(cherry_blossoms). Consider at least two functions to predict doy with temp. Compare them with PSIS or WAIC. Suppose March temperatures reach 9 degrees by the year 2050. What does your best model predict for the predictive distribution of the day-in-year that the cherry trees will blossom?  

**Question 4:** The data in data(Dinosaurs) are body mass estimates at different estimated ages for six different dinosaur species. See ?Dinosaurs for more details. Choose one or more of these species (at least one, but as many as you like) and model its growth. To be precise: Make a predictive model of body mass using age as a predictor. Consider two or more model types for the function relating age to body mass and score each using PSIS and WAIC. Which model do you think is best, on predictive grounds? On scientific grounds? If your answers to these questions differ, why? This is a challenging exercise, because the data are so scarce. But it is also a realistic example, because people publish Nature papers with even less data. So do your best, and I look forward to seeing your growth curves.

