---
title: ""
output: github_document
---

## Homework Week 1

**Question 1:** Suppose the globe tossing question (Chapter 2) had turned out to be 4 water and 11 land. Construct the posterior distribution, using grid approximation. Use the same flat prior as in the book.  

Response: To change the observations and get a new posterior based on a different dataset, adjust the likelihood

```{r}
# define grid 
p_grid <- seq(from = 0, to = 1, length.out = 20)
# define uniform prior 
prior <- rep(1,20)
# compute likelihood at each value in grid
likelihood <- dbinom(4, size = 4+11, prob = p_grid)
# compute likelihood of likelihood and prior 
unstd.posterior <- likelihood * prior
# standardize posterior 
posterior <- unstd.posterior/sum(unstd.posterior)
# plot 
plot(p_grid, posterior, type = "b", 
     xlab = "probability of water", ylab = "posterior probability")
# sample 
set.seed(100)
samples <- sample(p_grid, prob = posterior, size = 10000, replace = T)
```

**Question 2:** Now suppose the data are 4 water and 2 land. Compute the posterior again but this time use a prior that is zero below *p* = 0.5 and a constant above *p* = 0.5. This corresponds to prior information that a majority of the Earth's surface is water. 

```{r}
# define grid 
p_grid <- seq(from = 0, to = 1, length.out = 20)
# define new prior 
newprior <- ifelse(p_grid < 0.5, 0, 1)
# RICHARD's PRIOR - DISCUSS 
prior <- c(rep(0,500), rep(1,500))
# compute likelihood at each value in grid 
likelihood <- dbinom(4, size = 4+2, prob = p_grid)
# compute likelihood of likelihood and prior 
unstd.posterior <- likelihood * newprior
# standardize posterior 
posterior <- unstd.posterior/sum(unstd.posterior)
# plot 
plot(p_grid, posterior, type = "b", 
     xlab = "probability of water", ylab = "posterior probability")
# sample 
set.seed(100)
samples2 <- sample(p_grid, prob = posterior, size = 1e4, replace = T)
```

**Question 3:** For the posterior distribution from question 2, compute 89% percentile and HPDI intervals. Compare the widths of these intervals. Which is wider? Why? If you only had the information in the interval, what might you misunderstand about the shape of the posterior distribution?  


*Percentile Intervals:*
```{r, echo = F, warnings = F, message = F}
library(rethinking)
# define grid 
p_grid <- seq(from = 0, to = 1, length.out = 20)
# define new prior 
newprior <- ifelse(p_grid < 0.5, 0, 1)
# compute likelihood at each value in grid 
likelihood <- dbinom(4, size = 6, prob = p_grid)
# compute likelihood of likelihood and prior 
unstd.posterior <- likelihood * newprior
# standardize posterior 
posterior <- unstd.posterior/sum(unstd.posterior)
# sample
samples <- sample(p_grid , prob=posterior , size=1e4 , replace=TRUE )
# calculate percentile interval
PI(samples, prob = 0.89)
```


*HPDI Intervals:*
```{r, echo = F, warnings = F, message = F}
library(rethinking)
# define grid 
p_grid <- seq(from = 0, to = 1, length.out = 20)
# define new prior 
newprior <- ifelse(p_grid < 0.5, 0, 1)
# compute likelihood at each value in grid 
likelihood <- dbinom(4, size = 6, prob = p_grid)
# compute likelihood of likelihood and prior 
unstd.posterior <- likelihood * newprior
# standardize posterior 
posterior <- unstd.posterior/sum(unstd.posterior)
# sample
samples <- sample(p_grid , prob=posterior , size=1e4 , replace=TRUE )
# calculate HDPI interval 
HPDI(samples, prob = 0.89)
```

Response: The percentile interval is wider than the HPDI interval. This is because the percentile interval assigns the same probability mass to each **tail** whereas the HPDI interval calculates the narrowest interval possible containing the specified probability mass (0.89). The interval does not capture the distribution of values so by interpreting only the interval you may misinterpret the probability of certain parameter values.


**Question 4:** OPTIONAL CHALLENGE. Suppose there is bias in sampling so that Land is more likely than Water to be recorded. Specifically, assume that 1-in-5 (20%) of Water samples are accidentally recorded instead as "Land". First, write a generative simulation of this sampling process. Assuming the true proportion of Water is 0.70, what proportion does your simulation tend to produce instead? Second, using a simulated sample of 20 tosses, compute the unbiased posterior distribution of the true proportion of water.  

NOTE: Checked Richard's answers for this one.  

If there are 0.7/1 ways to sample water and 0.8/1 ways for water to be reported as water - there are 0.7 x 0.8 ways t observe water (garden multiplies independent events)


```{r}
N <- 1e5
set.seed(100)
W <- rbinom(N, size = 20, prob = 0.7*0.8)
mean(W/20) # biased expectation
```

This is the biased expectation. Now estimate posterior distribution using grid approximation, accounting for biased sampling rate.

```{r}
# simulate data 
W <- rbinom(1, size=20, prob=0.7*0.8)

# compute posterior 
grid_p <- seq(from=0, to=1, len=100)
pr_p <- dbeta(grid_p, 1, 1)
prW <- dbinom(W, 20, grid_p*0.8)
post <- prW*pr_p
```

Now calculate posterior ignoring bias and plot to compare.

```{r}
# simulate data 
W <- rbinom(1, size=20, prob=0.7*0.8)

# compute posterior 
grid_p <- seq(from=0, to=1, len=100)
pr_p <- dbeta(grid_p, 1, 1)
prW <- dbinom(W, 20, grid_p*0.8)
post <- prW*pr_p

# compute posterior ignoring bias
post_bad <- dbinom(W, 20, grid_p)

# plot
plot(grid_p, post, type = "l", lwd=4,
     xlab = "proportion water", ylab = "plausability")
lines(grid_p, post_bad, col = 2, lwd = 4)