---
title: ""
output: html_document
---

## Homework Week 1

**Question 1:** Suppose the globe tossing question (Chapter 2) had turned out to be 4 water and 11 land. Construct the posterior distribution, using grid approximation. Use the same flat prior as in the book.  

Response: To change the observations and get a new posterior based on a different dataset, adjust the likelihood

```{r}
# define grid 
p_grid <- seq(from = 0, to = 1, length.out = 20)
# define uniform prior 
prior <- rep(1,20)
# compute likelihood at each value in grid 
likelihood <- dbinom(4, size = 15, prob = p_grid)
# compute likelihood of likelihood and prior 
unstd.posterior <- likelihood * prior
# standardize posterior 
posterior <- unstd.posterior/sum(unstd.posterior)
# plot 
plot(p_grid, posterior, type = "b", 
     xlab = "probability of water", ylab = "posterior probability")
```

**Question 2:** Now suppose the data are 4 water and 2 land. Compute the posterior again but this time use a prior that is zero below *p* = 0.5 and a constant above *p* = 0.5. This corresponds to prior information that a majority of the Earth's surface is water. 

```{r}
# define grid 
p_grid <- seq(from = 0, to = 1, length.out = 20)
# define new prior 
newprior <- ifelse(p_grid < 0.5, 0, 1)
# compute likelihood at each value in grid 
likelihood <- dbinom(4, size = 6, prob = p_grid)
# compute likelihood of likelihood and prior 
unstd.posterior <- likelihood * newprior
# standardize posterior 
posterior <- unstd.posterior/sum(unstd.posterior)
# plot 
plot(p_grid, posterior, type = "b", 
     xlab = "probability of water", ylab = "posterior probability")
```

**Question 3:** For the posterior distribution from question 2, compute 89% percentile and HDPI intervals. Compare the widths of these intervals. Which is wider? Why? If you only had the information in the interval, what might you misunderstand about the shape of the posterior distribution?  


**Question 4:** OPTIONAL CHALLENGE. Suppose there is bias in sampling so that Land is more likely than Water to be recorded. Specifically, assume that 1-in-5 (20%) of Water samples are accidentally recorded instead of as "Land". First, write a generative simulation of this sampling process. Assuming the true proportion of Water is 0.70, what proportion does your simulation tend to produce instead? Second, using a simulated sample of 20 tosses, compute the unbiased posterior distribution of the true proportion of water.